---
title: "Exercises"
format: html
---

## Exercise A Segmentation
The goal is to classify each point of my trajectory as either:
static = TRUE or static = FLASE (animal is moving) 

### Preprocessing

```{r}
library(tidyverse)
library(jsonlite)
library(lubridate)
library(sf)

#import data

raw <- fromJSON("location-history.json", flatten = TRUE)

# preprocessing to convert to movement data
loc <- raw |>
  mutate(
    start_time = ymd_hms(startTime), #convert start-time into read date-time object
    end_time = ymd_hms(endTime),
    
    record_type = case_when(
      !is.na(activity.distanceMeters) ~ "activity",
      !is.na(visit.probability) ~ "visit",
      TRUE ~ "unknown" #If distanceMeters is NOT missing, assign activity,if visit probability exist, assign visit. 
    ),
    
    location = case_when(
      record_type == "activity" ~ activity.start,
      record_type == "visit" ~ visit.topCandidate.placeLocation
    ), # create a location column with activity start coordinate for activity and coordinates of visit if visit
    
    lat = as.numeric(str_extract(location, "(?<=geo:)[0-9.\\-]+")), #extract latuitude
    lon = as.numeric(str_extract(location, "(?<=,)[0-9.\\-]+")), # extract longitude
    
    activity_type = activity.topCandidate.type
  ) |>
  select(start_time, end_time, record_type, activity_type, lat, lon) |> #keep only useful columns
  filter(!is.na(lat), !is.na(lon)) |> #remove rows without coordinates
  arrange(start_time)  #sort by time

```


```{r}
#Convert coordinares to distances in meters

loc_sf <- loc |>
  st_as_sf(coords = c("lon", "lat"), crs = 4326) |>
  st_transform(2056)

# separate coordinates in 2 columns like the Sabi example.
coords <- st_coordinates(loc_sf)

loc_sf <- loc_sf |>
  mutate(
    E = coords[, 1],
    N = coords[, 2]
  )
```


### Task 1: Calculate distances

#### Step a) Specify a temporal window v

```{r}
# we choose 30 minstes before and after each points
v <- 30
```
#### Steb b) measure distance to every point within v

```{r}
loc_sf <- loc_sf |>
  rowwise() |> # do the calculation one row at a time
  mutate(
    stepMean = mean( #the average distance from the current point to nearby points in time.
      as.numeric(
        st_distance(
          geometry,
          loc_sf$geometry[
            abs(as.numeric(difftime(loc_sf$start_time, start_time, units = "mins"))) <= v &
              abs(as.numeric(difftime(loc_sf$start_time, start_time, units = "mins"))) > 0
          ]
        )
      ),
      na.rm = TRUE
    )
  ) |>
  ungroup()

summary(loc_sf$stepMean)
```

### Task 2 Specify and apply threshold d


```{r}
summary(loc_sf$stepMean)
hist(loc_sf$stepMean)
boxplot(loc_sf$stepMean)

#use mean of stepmean as threshold 
d <- mean(loc_sf$stepMean, na.rm = TRUE)

#create the static column 

loc_sf <- loc_sf |>
  mutate(static = stepMean < d)


```

### Task 3  Visualize segmented trajectory

```{r}
loc_sf |>
  ggplot(aes(E, N)) +
  geom_path() +
  geom_point(aes(colour = static)) +
  coord_equal() +
  theme_minimal()
```

### Task 4  Segment-based analysis

```{r}
rle_id <- function(vec) {
  x <- rle(vec)$lengths
  as.factor(rep(seq_along(x), times = x))
}

# Create Segement IDs 
loc_sf <- loc_sf |>
  mutate(segment_id = rle_id(static))



#visualise only moving segements

loc_sf |>
  filter(static == FALSE) |>
  ggplot(aes(E, N)) +
  geom_path(aes(colour = segment_id)) +
  geom_point(aes(colour = segment_id)) +
  coord_equal() +
  theme_minimal()


#Calculate duration per segment
segment_summary <- loc_sf |>
  st_drop_geometry() |>
  group_by(segment_id, static) |>
  summarise(
    start_time = min(start_time),
    end_time = max(end_time),
    duration_min = as.numeric(difftime(end_time, start_time, units = "mins")),
    .groups = "drop"
  )

segment_summary

#remonve segements with duration < 5 min 
valid_segments <- segment_summary |>
  filter(duration_min >= 5)

loc_filtered <- loc_sf |>
  filter(segment_id %in% valid_segments$segment_id)

# Final plot 
loc_filtered |>
  filter(static == FALSE) |>
  ggplot(aes(E, N)) +
  geom_path(aes(colour = segment_id)) +
  geom_point(aes(colour = segment_id)) +
  coord_equal() +
  theme_minimal()


```

## Exercise B Similarity

```{r}
library(SimilarityMeasures)

pedestrian <- read_delim("pedestrian.csv")
names(pedestrian)
```

### Task 1 similarity measures 

```{r}
#Visualize all trajectories
ggplot(pedestrian, aes(E, N, colour = as.factor(TrajID))) +
  geom_path() +
  geom_point() +
  coord_equal() +
  facet_wrap(~ TrajID, ncol = 3, labeller = label_both)+
  theme_minimal()
```


### Task 2 Calculate similarity 

```{r}
#create a matrix per trajectory 
traj1 <- pedestrian |>
  filter(TrajID == 1) |>
  select(E, N) |>
  as.matrix()

traj2 <- pedestrian |>
  filter(TrajID == 2) |>
  select(E, N) |>
  as.matrix()

traj3 <- pedestrian |>
  filter(TrajID == 3) |>
  select(E, N) |>
  as.matrix()

traj4 <- pedestrian |>
  filter(TrajID == 4) |>
  select(E, N) |>
  as.matrix()

traj5 <- pedestrian |>
  filter(TrajID == 5) |>
  select(E, N) |>
  as.matrix()

traj6 <- pedestrian |>
  filter(TrajID == 6) |>
  select(E, N) |>
  as.matrix()

#Calculate similarity measures

similarity_results <- tibble(
  comparison_trajectory = c(2, 3, 4, 5, 6),
  
  DTW = c(
    DTW(traj1, traj2),
    DTW(traj1, traj3),
    DTW(traj1, traj4),
    DTW(traj1, traj5),
    DTW(traj1, traj6)
  ),
  
  EditDist = c(
    EditDist(traj1, traj2),
    EditDist(traj1, traj3),
    EditDist(traj1, traj4),
    EditDist(traj1, traj5),
    EditDist(traj1, traj6)
  ),
  
  Frechet = c(
    Frechet(traj1, traj2),
    Frechet(traj1, traj3),
    Frechet(traj1, traj4),
    Frechet(traj1, traj5),
    Frechet(traj1, traj6)
  ),
  
  LCSS = c(
    LCSS(traj1, traj2, pointSpacing = 10, pointDistance = 10, errorMarg = 10),
    LCSS(traj1, traj3, pointSpacing = 10, pointDistance = 10, errorMarg = 10),
    LCSS(traj1, traj4, pointSpacing = 10, pointDistance = 10, errorMarg = 10),
    LCSS(traj1, traj5, pointSpacing = 10, pointDistance = 10, errorMarg = 10),
    LCSS(traj1, traj6, pointSpacing = 10, pointDistance = 10, errorMarg = 10)
  )
)

similarity_results

# in long format 
similarity_long <- similarity_results |>
  pivot_longer(
    cols = c(DTW, EditDist, Frechet, LCSS),
    names_to = "measure",
    values_to = "value"
  )

#plot 
ggplot(similarity_long, aes(x = factor(comparison_trajectory), y = value, fill = factor(comparison_trajectory))) +
  geom_col() +
  facet_wrap(~ measure, scales = "free_y", ncol = 2) +
  labs(
    title = "Computed similarities using different measures\nbetween trajectory 1 to all other trajectories",
    x = "Comparison trajectory",
    y = "Value"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none"
  )

```

